---
title: "Global variables"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{global-variables}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(pipload)
```

## Introduction

the PIP project makes use of many directory paths and ad-hoc information that
cannot be gathered endogenously from the PIP project. Since `pipload` is the
package in charge of accessing and loading all the data in PIP, it is also in
charge of loading into memery the ad-hoc variables that are used across PIP.

## Functions

the two main functions are `pip_create_globals()` and `add_gls_to_env()`. The
former is the main function that creates all the variables and returns them in a
list. The latter is a convenient function to standardize the use of the global
variables across all PIP processes.

## Create globals

### Basics

`pip_create_globals()` makes use of three arguments that work together:
`root_dir`, `out_dir`, and `vintage`. `root_dir` refers to the directory path
where the PIP data resides. Since the directory path of the PIP data cannot be
made publicly available, it is stored in the environment variable
`PIP_ROOT_DIR`, which hosted privately by the PIP Technical Team.

For the sake of sake of demostrations, let's use a temporal directory as it it
where the root directory. The default behavior of `pip_create_globals()` is to
create the the whole directory structure in case it does not exist:

```{r fake-root-1}
fake_dir1 <- paste0(tempdir(), "/1")
Sys.setenv(PIP_ROOT_DIR = fake_dir1)
gls <- pip_create_globals()
fs::dir_tree(fake_dir1, type = "directory")
```

In addition to create the global vartiables, `pip_create_globals()` creates two
main folders in case they don't exist, `PIP-Data_QA` and
`pip_ingestion_pipeline`. This is useful if the user does not want to work in
the main PIP data folder structure. Let's focus for now on the `pc_data`
sub-directory inside `pip_ingestion_pipeline`. This folder has two other
folders, `cache` and `output`. Within the latter you see a vintage folder
`r format(Sys.Date(), "%Y%m%d")` with the form `"%Y%m%d"`.

Since by default both input and output data are stored within the structure of
`root_dir` directory, the value of `out_dir` is the same as `root_dir`. Yet, the
user can specify a different directory path to store the output from the PIP
pipeline.

```{r fake-root-2}
fake_dir2 <- paste0(tempdir(), "/2")
fake_dir3 <- paste0(tempdir(), "/3")
Sys.setenv(PIP_ROOT_DIR = fake_dir2)
gls <- pip_create_globals(out_dir = fake_dir3)
fs::dir_tree(fake_dir2, type = "directory")
fs::dir_tree(fake_dir3, type = "directory")
```

In the case above, folder `fake_dir2` represents `root_dir` and `fake_dir3` is
the output directory, `out_dir`. You can see that folder
`pip_ingestion_pipeline` is available in both `fake_dir2` and `fake_dir3`, but
`output` was created only on `fake_dir3` and not in `fake_dir2`. It does not
mean that in reality `fake_dir2` does not contain an `output` folder, but that
given that we are working with temporal directories, the `output` folder does
not exist in either of them.

### Vintages

This is where the option `vintage` comes into play. This argument refers to the
name of the sub-directories inside the `output` folder. It can take two special
values, "latest" or "new", or any other character. If it is "latest" (default),
the most recent version available in the vintage directory of the form "%Y%m%d"
will be used. If no folder exists with this form, a new folder with the date of
the execution will be created. If it is "new", a new folder with a name of the
form "%Y%m%d" will be created. All the names will be coerced to lower cases.

So, let's pretend that, inside the official directory, `fake_dir1`, the most
recent vintage of the `output` sub-directories is
`r format(Sys.Date(), "%Y%m%d")` .Yet, you don't want to mess up with it, so you
want to create a new vintage, "temp_out", and you also want to do it in a
directory different to the official one, `fake_dir2`. In this, can do something
like this.

```{r vintage}
Sys.setenv(PIP_ROOT_DIR = fake_dir2)
gls <- pip_create_globals(root_dir = fake_dir1, 
                          out_dir  = fake_dir2, 
                          vintage  = "temp_out")
fs::dir_tree(fake_dir1, type = "directory")
fs::dir_tree(fake_dir2, type = "directory")
```

As you can see, directory `temp_out` was created inside `output` of the
`fake_dir2` directory.
