---
title: "Read and Write with `pins`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Read and Write with `pins`}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
devtools::load_all(".")
# library(pipload)
```

## Quick Start: Reading and Writing with `pins`

This vignette shows how to use the `pip_write` and `pip_read` functions to save and retrieve data with versioning using the `pins` package. We'll use a temporary board for demonstration, but the same logic applies to your real boards.

## Create a temporary board

```{r}
board <- pins::board_temp(versioned = TRUE)
name <- "x"
```

## Write several versions of an object

```{r}
pip_write(board, x = 1:5,  name = name)
Sys.sleep(3)
pip_write(board, x = 1:10, name = name)
Sys.sleep(3)
pip_write(board, x = 4:9,  name = name)
```

## Read 

### the latest version

```{r}
pip_read(board, name)
```

### List all available versions

```{r}
vers <- pip_read(board, name, version = "available")
vers
```

### Read a specific version by version string

```{r}
version <- vers[3, version]
pip_read(board, name, version = version)
```

### Read previous versions by index

```{r}
pip_read(board, name, version = -1) # previous version
pip_read(board, name, version = -2) # two versions back
```
