---
title: "Read and Write with `pins`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Read and Write with `pins`}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
devtools::load_all(".")
# library(pipload)
```

## Quick Start: Reading and Writing with `pins`

This vignette shows how to use the `pip_write` and `pip_read` functions to save and retrieve data with versioning using the `pins` package. We'll use a temporary board for demonstration, but the same logic applies to your real boards.

## Create a temporary board

```{r}
board <- pins::board_temp(versioned = TRUE)
name <- "x"
```

## Write several versions of an object

```{r}
pip_write(board, x = 1:5,  pin_name = name)
Sys.sleep(3)
pip_write(board, x = 1:10, pin_name = name)
Sys.sleep(3)
pip_write(board, x = 4:9,  pin_name = name)
```

## Read 

### the latest version

```{r}
pip_read(board = board, 
         pin_name = name)
```

### List all available versions

```{r}
vers <- pip_read(board, 
                 pin_name = name, 
                 version = "available")
vers
```

### Read a specific version by version string

```{r}
version <- vers[3, version]
pip_read(board, pin_name = name, version = version)
```

### Read previous versions by index

```{r}
pip_read(board, pin_name =  name, version = -1) # previous version
pip_read(board,pin_name =  name, version = -2) # two versions back
```



## Build higher-level functions

`pip_read()` and `pip_write()` are the backbone of any other function in the `{pipload}` package. For instance, the function `load_dlw_data()` makes use of all the functionality of pins and `pip_read()`.

Below are some real-world examples of how you might use these higher-level functions in your workflow:

```{r load_dlw_data}
# Get the latest release and set up the working environment
lr <- pipfun::get_latest_pip_release()
pipfun::setup_working_release(release = lr$release,
                              identity = lr$identity,
                              verbose = FALSE)

# Using pin_name (with extension)
load_dlw_data(pin_name = "HRV_2011_EU-SILC_V01_M_V04_A_GMD_GPWG.qs") |> 
  head()

# Using pin_name (without extension also works)
load_dlw_data(pin_name = "HRV_2011_EU-SILC_V01_M_V04_A_GMD_GPWG") |> 
  head()

# Using country and year
load_dlw_data(country_code = "HRV", year = 2011) |> 
  head()

# Find all data for a country
find_dlw_data(country_code = "HRV") |> 
  head()

# Find the latest year in EACH module for a country
find_dlw_data(country_code = "HRV", latest_year = TRUE)
```


