---
title: "Read and Write Versioned Data with `pipload`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Read and Write Versioned Data with `pipload`}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(pipload)
```

## Read and write versioned data

`pip_read()` and `pip_write()` provide a small interface to the `{stamp}`
package for saving and retrieving versioned artifacts. A stamp alias identifies
the storage root. Initialize the alias once before using either function.

## Create a temporary stamp alias

The following setup creates an isolated storage root for this example. In a
project workflow, the alias is normally initialized by the project setup code.

```{r create-alias}
alias <- paste0("pipload_vignette_", Sys.getpid())
root <- fs::path_temp(alias)
fs::dir_create(root)
stamp::st_init(root, alias = alias)
name <- "x"
```

## Write several versions of an object

Pass the object as the first argument and identify it with `id`. The short
delays ensure that each write is recorded as a distinct version.

```{r write-versions}
pip_write(1:5, id = name, alias = alias)
Sys.sleep(1)
pip_write(1:10, id = name, alias = alias)
Sys.sleep(1)
pip_write(4:9, id = name, alias = alias)
```

## Read versions

### Read the latest version

With the default `version = NULL`, `pip_read()` loads the latest version.

```{r read-latest}
pip_read(id = name, alias = alias)
```

### List available versions

Use `version = "available"` to return version metadata. The `vintage` column
uses zero for the latest version, negative one for the previous version, and so
on.

```{r list-versions}
versions <- pip_read(
  id = name,
  alias = alias,
  version = "available"
)
versions
```

### Read a specific version

The version identifier from the metadata can be passed back to `pip_read()`.

```{r read-specific}
version_id <- versions[3, version_id]
pip_read(id = name, alias = alias, version = version_id)
```

### Read previous versions by index

Negative integer values select versions relative to the latest version.

```{r read-previous}
pip_read(id = name, alias = alias, version = -1)
pip_read(id = name, alias = alias, version = -2)
```

## Build higher-level functions

`pip_read()` and `pip_write()` are the backbone of higher-level functions in
the `{pipload}` package. For example, `load_dlw_data()` uses `pip_read()` to
load data from the configured DLW release.

The following examples require access to a configured PIP working release and
are therefore not evaluated when this vignette is built:

```{r load-dlw-data, eval = FALSE}
# Get the latest release and set up the working environment
lr <- pipfun::get_latest_pip_release()
pipfun::setup_working_release(
  release = lr$release,
  identity = lr$identity,
  verbose = FALSE
)

# Using id_name (with extension)
load_dlw_data(id_name = "HRV_2011_EU-SILC_V01_M_V04_A_GMD_GPWG.qs2") |>
  head()

# Using id_name without an extension also works
load_dlw_data(id_name = "HRV_2011_EU-SILC_V01_M_V04_A_GMD_GPWG") |>
  head()

# Using country and survey year
load_dlw_data(country_code = "HRV", surveyid_year = 2011) |>
  head()

# Find all data for a country
find_dlw_data(country_code = "HRV") |>
  head()

# Find the latest year in each module for a country
find_dlw_data(country_code = "HRV", latest_year = TRUE)
```
